Concurrent Programming with Harmony

### 2020-08-18 book/paper.tex
Latest

    process X:                          process Y:
        X0: flagX = True                   Y0: flagY = True
        X1: if not flagY goto X4           Y1: if not flagX goto Y4
        X2: flagX = False                  Y2: flagY = False
        X3: goto X0                        Y3: goto Y0
        X4: ...critical section...         Y4: ...critical section...
        X5: flagX = False                  Y5: flagY = False
\end{verbatim}
\end{code}
\item A colleague of the author asked if the first two assignments in
Peterson's algorithm (setting \textit{flags}[\textit{self}])
to \texttt{True} and \textit{turn} to 1 - \textit{self}) can be reversed.
After all, they are different variables assigned independent values---in a
sequential program one could surely swap the two assignments.
See if you can figure out for yourself if the two assignments can be
reversed.  Then run the program in \autoref{fig:peterson} after reversing
the two assignments and describe in English what happens.
\item Bonus question:
Can you generalize Peterson's algorithm to more than two processes?
\item Bonus question:
Implement
\href{https://en.wikipedia.org/wiki/Dekker%27s_algorithm}{Dekker's Algorithm},
\href{https://en.wikipedia.org/wiki/Szymanski%27s_algorithm}{Szymański's Algorithm}, or the
\href{https://en.wikipedia.org/wiki/Lamport%27s_bakery_algorithm}{Bakery Algorithm}.
Note that the last one uses unbounded state, so you should modify the processes so they
only try to enter the critical section a bounded number of times.
\end{problems}

@@

>>>>

    process X:                          process Y:
        X0: flagX = True                   Y0: flagY = True
        X1: if not flagY goto X4           Y1: if not flagX goto Y4
        X2: flagX = False                  Y2: flagY = False
        X3: goto X0                        Y3: goto Y0
        X4: ...critical section...         Y4: ...critical section...
        X5: flagX = False                  Y5: flagY = False
\end{verbatim}
\end{code}
\item A colleague of the author asked if the first two assignments in
Peterson's algorithm (setting \textit{flags}[\textit{self}])
to \texttt{True} and \textit{turn} to 1 - \textit{self}) can be reversed.
After all, they are different variables assigned independent values---in a
sequential program one could surely swap the two assignments.
See if you can figure out for yourself if the two assignments can be
reversed.  Then run the program in \autoref{fig:peterson} after reversing
the two assignments and describe in English what happens.
\item Bonus question:
Can you generalize Peterson's algorithm to more than two processes?
\item Bonus question:
Implement
\href{https://en.wikipedia.org/wiki/Dekker%27s_algorithm}{Dekker's Algorithm},
\href{https://en.wikipedia.org/wiki/Eisenberg_%26_McGuire_algorithm}{Eisenstein and McGuire's Algorithm},
\href{https://en.wikipedia.org/wiki/Szymanski%27s_algorithm}{Szymański's Algorithm}, or the
\href{https://en.wikipedia.org/wiki/Lamport%27s_bakery_algorithm}{Lamport's Bakery Algorithm}.
Note that the last one uses unbounded state, so you should modify the processes so they
only try to enter the critical section a bounded number of times.
\end{problems}

@@

Harmony does not have a \textbf{return} statement.  (You can assign a return value
to a method by setting the \textit{result} variable.)  Neither does it support
\textbf{break} or \textbf{continue} statements in loops.  One reason for their absence is
that, particularly in concurrent programming, such control flow directions are highly
error-prone.  It's easy to forget to, say, release a lock when returning a value in the
middle of a method, and a major source of bugs in practice.
